---
title: "Take-home exercise01" 
description: "Part1 and Part2" 
date: "May 11, 2025" 
date-modified: "JUL 11, 2025" 
format: html
author: "YUAN Yihao" 
editor: visual 
execute: 
  eval: true 
  echo: true 
  warning: false 
  freeze: true 
---

# Part 1

# 1.0 Overview

## 1.1 Background

Analysis on demographic structures and distribution of Singapore in 2024

## 1.2 Objective

Identify the gender distribution by age

Identify the age distribution by region

Conduct more detailed analysis on the regions with distinct characteristics

## 1.3 The Data

*Singapore Residents by Planning Area / Subzone, Single Year of Age and Sex, June 2024* dataset shares by [Department of Statistics, Singapore (DOS)](https://www.singstat.gov.sg/)

# 2. Packages

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

# 3. Getting Started

## 3.1 Importing data

```{r}
res_data <- read_csv("respopagesex2024/respopagesex2024.csv")
```

## 3.2 Data cleaning

### 3.2.1 Check missing values

```{r}
res_data %>% filter(if_any(everything(), is.na))
```

```{r}
colSums(is.na(res_data))
```

### 3.2.2 Change data values

```{r}
res_data <- res_data %>%
  mutate(Age = ifelse(Age == "90_and_Over", "90", Age),
         Age = as.numeric(Age)) 
```

### 3.2.3 Check and Convert data type

```{r}

str(res_data)


res_data <- res_data %>%
  mutate(
    Age = as.numeric(Age),
    Pop = as.numeric(Pop)
  )
```

### 3.2.4 Data glimpse

```{r}
str(res_data)
```

# 4. Exploratory Visual Analysis

## 4.1 Population by Age and Sex

The grey area represents the lower population for males and females. The colored bars represent the number of population exceeded another gender

```{r}
library(dplyr)
library(ggplot2)


grouped_data <- res_data %>%
  group_by(Age, Sex) %>%
  summarise(Pop = sum(Pop, na.rm = TRUE), .groups = "drop")


wide_data <- grouped_data %>%
  pivot_wider(names_from = Sex, values_from = Pop, values_fill = 0) %>%
  mutate(
    min_pop = pmin(Males, Females),
    male_extra = ifelse(Males > Females, Males - Females, 0),
    female_extra = ifelse(Females > Males, Females - Males, 0)
  )


ggplot() +

  geom_col(data = wide_data, aes(x = Age, y = min_pop), 
           fill = "gray70", width = 0.6) +


  geom_col(data = wide_data, aes(x = Age, y = male_extra), 
           fill = "steelblue", width = 0.6) +


  geom_col(data = wide_data, aes(x = Age, y = female_extra), 
           fill = "tomato", width = 0.6) +

  labs(
    title = "Population by Age and Sex",
    x = "Age",
    y = "Population"
  ) +
  theme_minimal()


```

Observations: Obviously, in the age range of 0 to 30, the male population is larger than the female population. However, from 30 to 90, the female population is significantly larger than the male population.

## 4.2 Population by Age Group and Planning Area Analyze the number of the aging population in each region.

```{r}
res_data_grouped <- res_data %>%
  mutate(AgeGroup = case_when(
    Age >= 0 & Age <= 20 ~ "Under Age",
    Age >= 21 & Age <= 60 ~ "Adult",
    Age >= 61 ~ "Aging"
  )) %>%
  group_by(PA, AgeGroup) %>%
  summarise(Pop = sum(Pop, na.rm = TRUE), .groups = "drop")

res_data_grouped_filtered <- res_data_grouped %>%
  filter(AgeGroup != "Adult")

ggplot(res_data_grouped_filtered, aes(x = Pop, y = reorder(PA, Pop), fill = AgeGroup)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.4) +
  labs(
    title = "Population by Age Group and Planning Area",
    x = "Population",
    y = "Planning Area",
    fill = "Age Group"
  )+ theme_minimal(base_size = 10) +
  theme(
    axis.text.y = element_text(size = 5, margin = margin(r = 5)),
    legend.position = "top",
    plot.margin = margin(10, 10, 10, 10)
  )



```

Observations: Given the large number of regions, it is proposed to construct new variables in order to more intuitively analyze and understand the distribution and proportion of the aging population.

## 4.3 Aging-to-Under-Age Population Ratio by Planning Area

```{r}
library(dplyr)


age_summary <- res_data_grouped %>%
  filter(AgeGroup %in% c("Under Age", "Aging")) %>%
  group_by(PA, AgeGroup) %>%
  summarise(Total = sum(Pop), .groups = "drop")


age_ratio <- age_summary %>%
  tidyr::pivot_wider(names_from = AgeGroup, values_from = Total) %>%
  mutate(ratio = Aging / `Under Age`)
```

```{r}
ggplot(age_ratio, aes(x = ratio, y = reorder(PA, ratio))) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Aging-to-Under-Age Population Ratio by Planning Area",
    x = "Aging / Under Age Ratio",
    y = "Planning Area"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 5),
    plot.title = element_text(face = "bold")
  )
```

Observation: From this chart, we can see the ranking of regions with a larger proportion of the elderly population.

# PART 2 Peer Learning

Comment for 2nd plot in https://pengxinhuang.netlify.app/take-home_ex01/#age-structure-comparison-across-planning-areas

pros
1. Appropriate type of plot. Easier to know the distribution.
2. Choose some typical area to represent diffirent area's distribution.
3. Put plots together to compare

cons
1. X-axis: The data has negative numbers
2. Y-axis: The text overlaps, and the last group is NA
3. Overall: The X-axis index of the four pictures are different, which may cause some population misguidance


```{r}
res_data <- res_data %>%
  mutate(Age_group = case_when(
    Age == "90_and_Over" ~ "90_and_Over",
    TRUE ~ paste0(floor(as.numeric(Age) / 10) * 10, "-", floor(as.numeric(Age) / 10) * 10 + 9)
  ))
agg_data <- res_data %>%
  group_by(PA, Age_group, Sex) %>%
  summarise(Pop = sum(Pop, na.rm = TRUE)) %>%
  ungroup()
agg_data <- agg_data %>%
  mutate(Pop_plot = ifelse(Sex == "Males", -Pop, Pop))
ggplot(agg_data %>% filter(PA %in% c("Bedok", "Bukit Timah", "Outram", "Tampines")),
       aes(x = Pop_plot, y = Age_group, fill = Sex)) +
  geom_bar(stat = "identity", width = 0.8) +
  scale_fill_manual(values = c("Males" = "steelblue", "Females" = "tomato")) +
  scale_x_continuous(labels = abs) +
  facet_wrap(~PA, ncol = 2, scales = "fixed") +
  labs(title = "Population Pyramids of Selected Planning Areas (2024)",
    subtitle = "Representing new township (Tampines), mature estate (Bedok), high-income area (Bukit Timah),\nand historical district (Outram)",
    x = "Population",
    y = NULL,
    fill = "Sex") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),
    plot.subtitle = element_text(size = 9, hjust = 0.5),
    strip.text = element_text(face = "bold")
  )


```
