---
title: "Take-home exercise01"
author: "YUAN yihao"
---

# 1. Overview

## 1.1 Background

Analysis on demographic structures and distribution of Singapore in 2024

## 1.2 Objective

Identify the gender distribution by age

Identify the age distribution by region

Conduct more detailed analysis on the regions with distinct characteristics

## 1.3 The Data

*Singapore Residents by Planning Area / Subzone, Single Year of Age and Sex, June 2024* dataset shares by [Department of Statistics, Singapore (DOS)](https://www.singstat.gov.sg/)

# 2. Packages

```{r}
pacman::p_load(ggiraph, plotly, 
               patchwork, DT, tidyverse) 
```

# 3. Getting Started

## 3.1 Importing data

```{r}
res_data <- read_csv("respopagesex2024/respopagesex2024.csv")
```

## 3.2 Data cleaning

### 3.2.1 Check missing values

```{r}
res_data %>% filter(if_any(everything(), is.na))
```

```{r}
colSums(is.na(res_data))
```

### 3.2.2 Change data values

```{r}
res_data <- res_data %>%
  mutate(Age = ifelse(Age == "90_and_Over", "90", Age),
         Age = as.numeric(Age)) 
```

### 3.2.3 Check and Convert data type

```{r}

str(res_data)


res_data <- res_data %>%
  mutate(
    Age = as.numeric(Age),
    Pop = as.numeric(Pop)
  )
```

### 3.2.4 Data glimpse

```{r}
str(res_data)
```

# 4. Exploratory Visual Analysis

## 4.1 Population by Age and Sex

The grey area represents the lower population for males and females. The colored bars represent the number of population exceeded another gender

```{r}
library(dplyr)
library(ggplot2)


grouped_data <- res_data %>%
  group_by(Age, Sex) %>%
  summarise(Pop = sum(Pop, na.rm = TRUE), .groups = "drop")


wide_data <- grouped_data %>%
  pivot_wider(names_from = Sex, values_from = Pop, values_fill = 0) %>%
  mutate(
    min_pop = pmin(Males, Females),
    male_extra = ifelse(Males > Females, Males - Females, 0),
    female_extra = ifelse(Females > Males, Females - Males, 0)
  )


ggplot() +

  geom_col(data = wide_data, aes(x = Age, y = min_pop), 
           fill = "gray70", width = 0.6) +


  geom_col(data = wide_data, aes(x = Age, y = male_extra), 
           fill = "steelblue", width = 0.6) +


  geom_col(data = wide_data, aes(x = Age, y = female_extra), 
           fill = "tomato", width = 0.6) +

  labs(
    title = "Population by Age and Sex",
    x = "Age",
    y = "Population"
  ) +
  theme_minimal()


```

Observations: Obviously, in the age range of 0 to 30, the male population is larger than the female population. However, from 30 to 90, the female population is significantly larger than the male population.

## 4.2 Population by Age Group and Planning Area Analyze the number of the aging population in each region.

```{r}
res_data_grouped <- res_data %>%
  mutate(AgeGroup = case_when(
    Age >= 0 & Age <= 20 ~ "Under Age",
    Age >= 21 & Age <= 60 ~ "Adult",
    Age >= 61 ~ "Aging"
  )) %>%
  group_by(PA, AgeGroup) %>%
  summarise(Pop = sum(Pop, na.rm = TRUE), .groups = "drop")

res_data_grouped_filtered <- res_data_grouped %>%
  filter(AgeGroup != "Adult")

ggplot(res_data_grouped_filtered, aes(x = Pop, y = reorder(PA, Pop), fill = AgeGroup)) +
  geom_col(position = position_dodge(width = 0.9), width = 0.4) +
  labs(
    title = "Population by Age Group and Planning Area",
    x = "Population",
    y = "Planning Area",
    fill = "Age Group"
  ) +
  scale_y_discrete(expand = expansion(mult = c(0.4, 0.4))) + 
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 3, margin = margin(r = 5)), 
    legend.position = "top"
  )


```

Observations: Given the large number of regions, it is proposed to construct new variables in order to more intuitively analyze and understand the distribution and proportion of the aging population.

## 4.3 Aging-to-Under-Age Population Ratio by Planning Area

```{r}
library(dplyr)


age_summary <- res_data_grouped %>%
  filter(AgeGroup %in% c("Under Age", "Aging")) %>%
  group_by(PA, AgeGroup) %>%
  summarise(Total = sum(Pop), .groups = "drop")


age_ratio <- age_summary %>%
  tidyr::pivot_wider(names_from = AgeGroup, values_from = Total) %>%
  mutate(ratio = Aging / `Under Age`)
```

```{r}
ggplot(age_ratio, aes(x = ratio, y = reorder(PA, ratio))) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Aging-to-Under-Age Population Ratio by Planning Area",
    x = "Aging / Under Age Ratio",
    y = "Planning Area"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 5),
    plot.title = element_text(face = "bold")
  )
```

Observation: From this chart, we can see the ranking of regions with a larger proportion of the elderly population.
